---
title: "P8130 Final Project"
author: "Pengyuan Su (ps3195), Shuhong Xiang (sx2289), Yali Zhai (yz3959), Zhixing Wu (zw2709)"
date: "12/6/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(purrr)
library(olsrr)
library(dvmisc)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


```{r load_libraries}
library(tidyverse)
library(modelr)
library(arsenal)
library(broom)
library(modelr)
library(patchwork)
library(mgcv)
library(faraway)
library(skimr)
```

# Data Analysis


## Tidy:

`rate`: hate_crimes_per_100k_splc

`med_income`: median_household_income

`high_degree`: perc_population_with_high_school_degree

`non_citizen`: perc_non_citizen

`non_white`: perc_non_white


```{r import data}
hc_df = 
  read_csv(here::here("data/HateCrimes.csv")) %>% 
  janitor::clean_names() %>% 
  mutate(hate_crimes_per_100k_splc = as.numeric(hate_crimes_per_100k_splc)) %>% 
  drop_na() %>% 
  rename(., 
         rate = hate_crimes_per_100k_splc, 
         med_income = median_household_income, 
         high_degree = perc_population_with_high_school_degree, 
         non_citizen = perc_non_citizen, 
         non_white = perc_non_white)

head(hc_df)
```

### Description by table:
```{r}
char_df = hc_df %>% select(state, unemployment,urbanization)

skim(char_df) %>% knitr::kable(caption = "Table 1. Summry of State and Unemployment")
```

```{r}
num_df = hc_df %>% select(med_income, high_degree, non_white, gini_index, non_citizen, rate)

skim(num_df) %>% knitr::kable(caption = "Table 2. Summry of Numeric Variables")
```


### Change "unemployment" and "urbanization" from characters to corresponding numbers "1","2"

use 1 and 2 to denote the level low and high

```{r}
hc_df = 
  hc_df %>% 
  mutate(
    unemployment = case_when(
      unemployment == "low" ~ 1,
      unemployment == "high" ~ 2)
  ) %>% 
  mutate(
    urbanization = case_when(
      urbanization == "low" ~ 1,
      urbanization == "high" ~ 2)
  ) %>% 
  
  select(-state)
```


## Plot distribution of hate crimes rate

```{r plot distribution of rate}
hc_df %>% 
  ggplot(aes(x = rate, y = ..density..)) +
  geom_histogram(fill = "blue", alpha = .4) +
  geom_density(aes( x = rate, y = ..density..)) +
  theme_bw() +
  labs(title = "Figure1. Distribution for hate crimes per 100k",
       x = "hate crimes per 100k splc",
       y = " Count") +
  theme(plot.title = element_text(hjust = .5 ))
```


```{r plot distribution of rate 2}
hc_df_log = 
  hc_df %>% 
  mutate(lg = log(rate)) %>% 
  select(-rate)

hc_df_log%>% 
  ggplot(aes(x = lg, y = ..density..)) +
  geom_histogram(fill = "blue", alpha = .4, binwidth = 0.2) +
  geom_density(aes( lg, y = ..density..)) +
  theme_bw() +
  labs(title = "Figure2. Distribution for Log ",
       x = "Log (hate crimes per 100k splc)",
       y = " Count") +
  theme(plot.title = element_text(hjust = .5 ))


```

```{r}

# Create a quantile-quantile plot (QQplot)
qqnorm(hc_df_log$lg, col=2, pch=19, cex=1.5, main = "Figure3. Normal Q-Q Plot")

# Add a straight line which passes through the first and third quartiles.
qqline(hc_df_log$lg, col = 1,lwd=2,lty=2)
```


Based on the QQ plot in terms of the log(hate crimes per 100k splc), the points are nearly distributed along the straight line, so there is not severe departure from the normality.

```{r}
# Perform Shapiro-Wilk test
shapiro.test(hc_df_log$lg) 
```



H0: the log(hate crimes per 100k splc) is normally distributed
H1: the log(hate crimes per 100k splc) is not normally distributed

The W test statistic is 0.983 with p value of 0.7453.

Since the p value is greater than 0.05, so we fail to reject the null and we could conclude that the log(hate crimes per 100k splc) seems to be normally distributed under the significance level of 5%.



## Correlation

```{r }

hc_df %>% cor() %>% round(., 2) %>% knitr::kable(caption = "Table3. Correlation Matrix")
```


## Verify association

**income**

```{r}
income = 
 lm(rate ~ med_income, data = hc_df)
summary(income)
```


**high school degree**

```{r}
degree = 
 lm(rate ~ high_degree, data = hc_df)

summary(degree)
```


**non citizen**

```{r}
noncitizen = 
 lm(rate ~ non_citizen, data = hc_df)

summary(noncitizen)
```


**non white**

```{r}
nonwhite = 
 lm(rate ~ non_white, data = hc_df)

summary(nonwhite)
```


**gini index**

```{r}
gini = 
 lm(rate ~ gini_index, data = hc_df)

summary(gini)
```


**unemployment**

```{r}
unemployment = 
 lm(rate ~ unemployment, data = hc_df)

summary(unemployment)
```



**urbanization**

```{r}
urbanization = 
 lm(rate ~ urbanization, data = hc_df)

summary(urbanization)
```

We conduct the linear regression between Y and all variables X individuals, and the variable "med_income" and "gini_index" seem to be significant with p value smaller than 0.05. 


Since "med_income" and "gini_index" are significant predictors, then we use these two predictors as the basis of the model to explore further associations.



## Use "gini_index" as the Model Basis

```{r}
# step1
fit1 <- lm(rate ~ unemployment, data = hc_df)
tidy(fit1)
fit2 <- lm(rate ~ urbanization, data = hc_df)
tidy(fit2)
fit3 <- lm(rate ~ med_income, data = hc_df)
tidy(fit3)
fit4 <- lm(rate ~ high_degree, data = hc_df)
tidy(fit4)
fit5 <- lm(rate ~ non_citizen, data = hc_df)
tidy(fit5)
fit6 <- lm(rate ~ gini_index, data = hc_df)
tidy(fit6)
fit7 <- lm(rate ~ non_white, data = hc_df)
tidy(fit7)

forward1 = lm(rate ~ gini_index, data = hc_df)

# step 2
fit1 <- update(forward1, . ~ . +unemployment)
tidy(fit1)
fit2 <- update(forward1, . ~ . +urbanization)
tidy(fit2)
fit3 <- update(forward1, . ~ . +med_income)
tidy(fit3)
fit4 <- update(forward1, . ~ . +high_degree)
tidy(fit4)
fit5 <- update(forward1, . ~ . +non_citizen)
tidy(fit5)
fit6 <- update(forward1, . ~ . +non_white)
tidy(fit6)

forward2 = update(forward1, . ~ . +high_degree)

#step 3
fit1 <- update(forward2, . ~ . +unemployment)
tidy(fit1)
fit2 <- update(forward2, . ~ . +urbanization)
tidy(fit2)
fit3 <- update(forward2, . ~ . +med_income)
tidy(fit3)
fit4 <- update(forward2, . ~ . +non_citizen)
tidy(fit4)
fit5 <- update(forward2, . ~ . +non_white)
tidy(fit5)


```

Based on the results, the significant predictors are "gini_index" and "high_degree".


## Use "med_income" as the Model Basis


```{r}
income_reg = lm(rate ~ med_income, data = hc_df)
# step 1
fit1 <- update(income_reg, . ~ . +unemployment)
tidy(fit1)
fit2 <- update(income_reg, . ~ . +urbanization)
tidy(fit2)
fit3 <- update(income_reg, . ~ . +gini_index)
tidy(fit3)
fit4 <- update(income_reg, . ~ . +high_degree)
tidy(fit4)
fit5 <- update(income_reg, . ~ . +non_citizen)
tidy(fit5)
fit6 <- update(income_reg, . ~ . +non_white)
tidy(fit6) 

forward1 = update(income_reg, . ~ . +gini_index)

#step 2
fit1 <- update(forward1, . ~ . +unemployment)
tidy(fit1)
fit2 <- update(forward1, . ~ . +urbanization)
tidy(fit2)
fit3 <- update(forward1, . ~ . +high_degree)
tidy(fit3)
fit4 <- update(forward1, . ~ . +non_citizen)
tidy(fit4)
fit5 <- update(forward1, . ~ . +non_white)
tidy(fit5)

#delete med_income, add high_degree
forward2 = lm(rate~gini_index + high_degree, data = hc_df)

#step 3
fit1 <- update(forward2, . ~ . +unemployment)
tidy(fit1)
fit2 <- update(forward2, . ~ . +urbanization)
tidy(fit2)
fit3 <- update(forward2, . ~ . +non_white)
tidy(fit3)
fit4 <- update(forward2, . ~ . +non_citizen)
tidy(fit4)


```

Based on the results, the significant predictors are "gini_index" and "high_degree".


## Stepwise Method

```{r}

mult.fit = lm(rate ~ ., data = hc_df)
step(mult.fit, direction='backward')

```
The stepwise method shows the relationship； rate ~ high_degree + gini_index.

The results are similar to the previous model, showing the "high_degree" and "gini_index" are significant.



# Interaction check

urbanization & high_degree

```{r}
##Urbanization -- high_degree
hc_df %>% 
  ggplot(aes(x = high_degree, y = rate, color = factor(urbanization))) +
  geom_point() +
  stat_smooth(method = "lm", se = F) +
  labs(title = "Figure4. high school degree vs. hate rate (By urbanization)",
       x = "percentage of adults (>25 yrs.) with a high school degree",
       y = " hate rate") +
  theme(plot.title = element_text(hjust = .5 ))

```


urbanization & gini_index

```{r}
##Urbanization -- gini
hc_df %>% 
  ggplot(aes(x = gini_index, y = rate, color = factor(urbanization))) +
  geom_point() +
  stat_smooth(method = "lm", se = F) +
  labs(title = "Figure5. gini index vs. hate rate (By urbanization)",
       x = "gini index",
       y = " hate rate") +
  theme(plot.title = element_text(hjust = .5 ))


```



unemployment & high_degree

```{r}
##Unemployment -- high_degree
hc_df %>% 
  ggplot(aes(x = high_degree, y = rate, color = factor(unemployment))) +
  geom_point() +
  stat_smooth(method = "lm", se = F) +
  labs(title = "Figure6. high school degree vs. hate rate (By unemployment)",
       x = "percentage of adults (>25 yrs.) with a high school degree",
       y = " hate rate") +
  theme(plot.title = element_text(hjust = .5 ))


```

unemployment & gini
```{r}
##Unemployment -- gini
hc_df %>% 
  ggplot(aes(x = gini_index, y = rate, color = factor(unemployment))) +
  geom_point() +
  stat_smooth(method = "lm", se = F) +
  labs(title = "Figure7. gini index vs. hate rate (By unemployment)",
       x = "gini index",
       y = " hate rate") +
  theme(plot.title = element_text(hjust = .5 ))

```


We check the interaction for `urbanization` and `unemployment` among `high_degree`, and `gini_index` and find that the two variables have interactions with urbanization and unemployment. Next step, we will check whether the interaction term is significant.


# Add Interaction to Model and Check Significance

Initial:

```{r}
model = 
  hc_df %>% 
  lm(rate ~ high_degree + gini_index, data = .)

summary(model)

```

Add Interaction: 

<<<<<<< HEAD
```{r}
model1 = 
  hc_df %>% 
  lm(rate ~ high_degree + gini_index + high_degree * gini_index, data = .)

summary(model1)
```


1. check for adding all interactions for the two models
=======
1. Check for Adding All Interactions for the Two Models
>>>>>>> be40ce3d9f4eefeab09cffaef6b17ce13f94601b

```{r}
inter_model_1= 
  hc_df %>% 
  lm(rate ~ high_degree * gini_index * urbanization , data = .)

summary(inter_model_1)
```

```{r}
inter_model_2 = 
  hc_df %>% 
  lm(rate ~ high_degree * gini_index * unemployment, data = .)

summary(inter_model_2)
```


Conclude: we find that the interaction terms are not significant since the p values are larger than 0.05, so we will not add them into the models.


So the final model is rate ~ high_degree + gini_index


# Model Diagnostic

```{r}
model = lm(rate ~ high_degree + gini_index, data = hc_df)

model_full = lm(rate ~ ., data = hc_df)
```

```{r}
cp = ols_mallows_cp(model, model_full)
cp

mse = get_mse(model, var.estimate = FALSE)
mse

model_summ = summary(model)
model_summ$adj.r.squared

model_full_summ = summary(model_full)
model_full_summ$adj.r.squared
```

## Cross-Validation Plot and Diagnostic Plots: 

```{r}
set.seed(1)
cv_df =
  crossv_mc(hc_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(
    linear_mod  = map(train, ~lm(rate ~ gini_index + high_degree, data = .x))) %>% 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, ~rmse(model = .x, data = .y)))


cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() + labs(title = "Figure8. RMSE Violin Graph from Cross-Validation", x = "Our model: rate ~ gini_index + high_degree")

mean(cv_df$rmse_linear)
```


```{r}
par(mfrow=c(2,2))
plot(model)
title("Figure9. Diagnostic Graphs", line = -1, outer = TRUE)
```


